### 2021-04-30 parts/02-abstract.tex
Added final report files

% A B S T R A C T
% ---------------

\begin{center}\textbf{Abstract}\end{center}

%This paper proposes a strategy for testing and comparing three control algorithms, LQR, LQG, and ADP, to control two two-DOF helicopters from a mobile device.  We will be using Raspberry Pi 3's as terminals for the wireless communication and MATLAB as our primary coding language.

Smart building energy management and/or monitoring has recently emerged as a
promising task due to global climate change issues. Such a task has received thorough attention
due to the advent of emerging Internet-of-Things (IoT) technology. Numerous building energy
management systems are commercially available in the market. Most of them lack modularity
or are driven by an overwhelming degree of manual configurations for performing building
automation. In this project, we develop a modular IoT-based framework for a smart building
energy management system (BEMS). The proposed framework is open-source and leverages the
features of decentralized multi-agent systems. The principal component of the framework is the
BEMS core that implements building automation features, such as an optimal scheduler,
displaying weather information, detecting new IoT devices, monitoring energy consumption,
and remotely controlling operation of IoT devices.
\cleardoublepage
%\newpage

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../finalReportMainV1"
%%% End: