### 2021-04-30 parts/03-acknowledgements.tex

% A C K N O W L E D G E M E N T S
% -------------------------------

\begin{center}\textbf{Acknowledgements}\end{center}

    \par Thanks to Dr. Suruz Miah for his advisement and preliminary work on this project.
    \par Thanks to Mr. Christopher Mattus for his assistance in setting up our lab environment.
    \par Thank you to everyone else who helped make this project possible and successful.

\cleardoublepage
%\newpage

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../finalReportMainV1"
%%% End: